---
title: "P8105 HW5"
author: "Jennifer Lee (UNI: jl2361)"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
echo = TRUE,
warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%")

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

## Data load & tidying
The code chunk below imports the data in individual spreadsheets contained in `./data/zip_data/`. To do this, I create a dataframe that includes the list of all files in that directory and the complete path to each file. As a next step, I `map` over paths and import data using the `read_csv` function. Finally, I `unnest` the result of `map`.
```{r}
full_df = 
  tibble(
    files = list.files("data/zip_data/"),
    path = str_c("data/zip_data/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest()
```

The result of the previous code chunk isn't tidy -- data are wide rather than long, and some important variables are included as parts of others. The code chunk below tides the data using string manipulations on the file, converting from wide to long, and selecting relevant variables. 
```{r}
tidy_df = 
  full_df %>% 
  mutate(
    files = str_replace(files, ".csv", ""),
    group = str_sub(files, 1, 3)) %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    values_to = "outcome",
    names_prefix = "week_") %>% 
  mutate(week = as.numeric(week)) %>% 
  select(group, subj = files, week, outcome)
```

## Plot
Finally, the code chunk below creates a plot showing individual data, faceted by group. 
```{r}
tidy_df %>% 
  ggplot(aes(x = week, y = outcome, group = subj, color = group)) + 
  geom_point() + 
  geom_path() + 
  facet_grid(~group)
```

This plot suggests high within-subject correlation -- subjects who start above average end up above average, and those that start below average end up below average. Subjects in the control group generally don't change over time, but those in the experiment group increase their outcome in a roughly linear way. 

# Problem 2

## Data load & tidying
First, we load the homicide data and do some data tidying. 
```{r}
homicide_df = 
  read_csv("data/homicide.csv") %>%
  janitor::clean_names() %>%
  mutate(
    state = toupper(state),
    city_state = str_c(city, state, sep = ", ") %>% 
    str_replace("Tulsa, AL", "Tulsa, OK"))
```

The raw data reports data on homicides in 50 U.S. cities describing the date, location, victim details, and case disposition. There are `r nrow(homicide_df)` rows/observations and `r ncol(homicide_df)` columns/variables after we create the `city_state` variable. We note that some state abbreviations are not completely capitalized and that there appears to be a typo for the city of Tulsa, so we fix these data entry errors in the above code chunk. 

Next, we create a dataframe summarizing the number of total homicides by city and state, in addition to the number of unsolved homicides categorized by "Closed without arrest" or "Open/No arrest."
```{r}
number_homicide_df = 
  homicide_df %>%
  group_by(city_state) %>%
  summarize(number_homicide = n()) 

number_unsolved_homicide_df = 
  homicide_df %>%
  filter(disposition %in% c("Closed without arrest", "Open/No arrest")) %>%
  group_by(city_state) %>%
  summarize(number_unsolved_homicide = n()) 

homicide = 
  left_join(number_unsolved_homicide_df, number_homicide_df, by = "city_state") 

homicide %>%
  knitr::kable()
```

## Using prop.test
We then create a `myproptest` function to apply the prop.test function to estimate the proportion of homicides that are unsolved, apply the broom::tidy, and save the output (estimated proportion and upper and lower bounds of the 95% CI). We use `myproptest` for the city of Baltimore, MD and make use of the `pull` function.
```{r}
myproptest = function(number_unsolved_homicide, number_homicide) {
  
    prop.test(number_unsolved_homicide, number_homicide, p = 0.5, alternative = "two.sided") %>%
    broom::tidy() %>%
    select(estimate, conf.low, conf.high)
  
}

baltimore = 
  homicide %>% 
    filter(city_state %in% c("Baltimore, MD"))

baltimore_results =
  myproptest(pull(baltimore, number_unsolved_homicide), pull(baltimore, number_homicide)) %>%
  rename(
    conf_low = conf.low,
    conf_high = conf.high)

baltimore_results
```

Now, we run `prop.test` for each of the cities in the dataset and extract both the proportion of unsolved homicides and the CIs for each within a “tidy” pipeline, making use of purrr::map2, list columns, and unnest.
```{r}
homicide_prop = 
  homicide %>%
   mutate(
    prop_data = map2(number_unsolved_homicide, number_homicide, ~ prop.test(.x, .y) %>%
      broom::tidy())) %>%
  unnest(prop_data) %>%
  select(city_state, estimate, conf.low, conf.high) %>%
  rename(
    conf_low = conf.low,
    conf_high = conf.high)

homicide_prop
```

## Plot
Finally, we create a plot that shows the estimates and CIs for each city, and organize cities according to the proportion of unsolved homicides. We see that Chicago, IL, New Orleans, LA, and Baltimore, MD are the top 3 cities with the highest proportions of unsolved homicides. 
```{r}
homicide_prop %>%
  mutate(
    city_state = fct_reorder(city_state, estimate, .desc = TRUE)) %>%
  ggplot(
    aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

# Problem 3

## Simulations
First, we set the following design elements:

Fix n=30
Fix σ=5
Set μ=0. 

We generate 5000 datasets from the model using `expand_grid` and call this `sim_df`:

x∼Normal[μ,σ]

In addition, we create a `sim_ttest` function to map to `sim_df` and save μ̂  and the p-value arising from a test of H:μ=0 using α=0.05
```{r}
sim_ttest = function(n = 30, mean, sigma = 5) {
  
  sim_data = tibble(
    x = rnorm(n = 30, mean, sd = sigma),
  )
    sim_data %>% 
    t.test(mu = 0, conf.level = 0.95) %>%
    broom::tidy() %>%
    select(estimate, p.value)

}

sim_df = 
  expand_grid(
    mean = 0,
    iter = 1:5000) %>% 
  mutate(
    ttest_df = map(.x = mean, ~ sim_ttest(mean = .x))) %>% 
  unnest(ttest_df) %>%
  rename(
      mu_hat = estimate,
      p_value = p.value)

sim_df
```

We repeat the above for μ={1,2,3,4,5,6}.
```{r}
sim_df2 = 
  expand_grid(
    mean = c(1, 2, 3, 4, 5, 6),
    iter = 1:5000) %>% 
  mutate(
    ttest_df = map(.x = mean, ~ sim_ttest(mean = .x))) %>% 
  unnest(ttest_df) %>%
  rename(
      mu_hat = estimate,
      p_value = p.value)

sim_df2
```

## Plots
Here is a plot showing the proportion of times the null was rejected (the power of the test assuming a 5% level of significance) on the y axis and the true value of μ on the x axis. As the effect size gets bigger, we see that power increases as well. 

```{r}
sim_df2 %>% 
  mutate(
    reject = if_else(p_value < 0.05, "y", "n")) %>%
  group_by(mean, reject) %>%
  summarize(
    num_reject = n()) %>%
  mutate(
    prop_reject = num_reject/5000) %>%
  filter(reject == "y") %>%
  ggplot(
    aes(x = mean, y = prop_reject)) +
  geom_point() + geom_line()
```

Next, here is a plot showing the average estimate of μ̂  on the y axis and the true value of μ on the x axis. There is also a second plot (right) showing the average estimate of μ̂  only in samples for which the null was rejected (assuming a 5% level of significance) on the y axis and the true value of μ on the x axis. 

The sample average of μ̂ across test for which the null is rejected is approximately equal to the true value of μ. As the error variance decreases, power increases.
```{r}
mean_plot1 =
  sim_df2 %>% 
  ggplot(
    aes(x = mean, y = mu_hat)) +
  geom_point(alpha = 0.5) +
  labs(title = "All samples")

mean_plot2 =
  sim_df2 %>% 
    mutate(
    reject = if_else(p_value < 0.05, "y", "n")) %>%
  filter(reject == "y") %>%
  ggplot(
    aes(x = mean, y = mu_hat)) +
  geom_point(alpha = 0.5) +
  labs(title = "Samples for null rejected")

mean_plot1 + mean_plot2
```
