---
title: "P8105 HW5"
author: "Jennifer Lee (UNI: jl2361)"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(patchwork)

knitr::opts_chunk$set(
echo = TRUE,
warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%")

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1
This zip file contains data from a longitudinal study that included a control arm and an experimental arm. Data for each participant is included in a separate file, and file names include the subject ID and arm.

Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time:

Start with a dataframe containing all file names; the list.files function will help
Iterate over file names and read in data for each subject using purrr::map and saving the result as a new variable in the dataframe
Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary
Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.

# Problem 2

## Data load & tidying
First, we load the homicide data and do some data tidying. 
```{r}
homicide_df = 
  read.csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") %>%
  janitor::clean_names() %>%
  mutate(
    state = toupper(state),
    city_state = str_c(city, state, sep = ", ") %>% 
    str_replace("Tulsa, AL", "Tulsa, OK"))
```

The raw data reports data on homicides in 50 U.S. cities describing the date, location, victim details, and case disposition. There are `r nrow(homicide_df)` rows/observations and `r ncol(homicide_df)` columns/variables after we create the `city_state` variable. We note that some state abbreviations were not completely capitalized and that there appears to be a typo for the city of Tulsa, so we fix these data entry errors. 

Next, we create a dataframe summarizing the number of total homicides by city and state, in addition to the number of unsolved homicides categroized by "Closed without arrest" or "Open/No arrest."
```{r}
number_homicide_df = 
  homicide_df %>%
  group_by(city_state) %>%
  summarize(number_homicide = n()) 

number_unsolved_homicide_df = 
  homicide_df %>%
  filter(disposition %in% c("Closed without arrest", "Open/No arrest")) %>%
  group_by(city_state) %>%
  summarize(number_unsolved_homicide = n()) 

homicide = 
  left_join(number_unsolved_homicide_df, number_homicide_df, by = "city_state") 

homicide %>%
  knitr::kable()
```

## Using prop.test
We then create a `myproptest` function to apply the prop.test function to estimate the proportio of homicides that are unsolved, apply the broom::tidy, and save the output (estimated proportion and upper and lower bounds of the 95% confidence interval) as a tibble. We use `myproptest` for the city of Baltimore, MD and note that the filtered dataframe needs to be converted to a list of vectors prior to applying `myproptest`.
```{r}
myproptest = function(x) {
  
    estimate = 
      prop.test(x[[2]], x[[3]]) %>% 
      broom::tidy() %>%
      select(estimate)
    conf_low = 
      prop.test(x[[2]], x[[3]]) %>% 
      broom::tidy() %>%
      select(conf.low)
    conf_high = 
      prop.test(x[[2]], x[[3]]) %>% 
      broom::tidy() %>%
      select(conf.high)

tibble(
  estimate, 
  conf_low,
  conf_high) %>%
  rename(
      conf_low = conf.low,
      conf_high = conf.high)

  }

baltimore = 
  homicide %>% 
    filter(city_state %in% c("Baltimore, MD")) %>% 
  as.list() %>%
  myproptest()

baltimore
```

Now, we run `prop.test` for each of the cities in the dataset and extract both the proportion of unsolved homicides and the confidence interval for each within a “tidy” pipeline, making use of purrr::map2, list columns, and unnest.
```{r}
homicide_prop = 
  homicide %>%
   mutate(
    prop_data = map2(number_unsolved_homicide, number_homicide, ~ prop.test(.x, .y) %>%
      broom::tidy())) %>%
  unnest(prop_data) %>%
  select(city_state, estimate, conf.low, conf.high) %>%
  rename(
    conf_low = conf.low,
    conf_high = conf.high
  )

homicide_prop
```

## Plot
Finally, we create a plot that shows the estimates and CIs for each city, and organize cities according to the proportion of unsolved homicides. We see that Chicago, IL, New Orleans, LA, and Baltimore, MD are the top 3 cities with the highest proportions of unsolved homicides. 
```{r}
homicide_prop %>%
  mutate(
    city_state = fct_reorder(city_state, estimate, .desc = TRUE)) %>%
  ggplot(
    aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

# Problem 3
When designing an experiment or analysis, a common question is whether it is likely that a true effect will be detected – put differently, whether a false null hypothesis will be rejected. The probability that a false null hypothesis is rejected is referred to as power, and it depends on several factors, including: the sample size; the effect size; and the error variance. In this problem, you will conduct a simulation to explore power in a one-sample t-test.

First set the following design elements:

Fix n=30
Fix σ=5
Set μ=0. Generate 5000 datasets from the model

x∼Normal[μ,σ]

For each dataset, save μ̂  and the p-value arising from a test of H:μ=0 using α=0.05. Hint: to obtain the estimate and p-value, use broom::tidy to clean the output of t.test.

```{r}
sim_ttest = function(n = 30, mean, sigma = 5) {
  
  sim_data = tibble(
    x = rnorm(n = 30, mean, sd = sigma),
  )
  mu_hat = 
    sim_data %>% 
    t.test() %>%
    broom::tidy() %>%
    select(estimate)
  p_value =
    sim_data %>% 
    t.test() %>%
    broom::tidy() %>%
    select(p.value)
  
  tibble(
    mu_hat, 
    p_value) %>%
    rename(
      mu_hat = estimate,
      p_value = p.value
    )

}

sim_df = 
  expand_grid(
    mean = 0,
    iter = 1:5000
  ) %>% 
  mutate(
    ttest_df = map(.x = mean, ~ sim_ttest(mean = .x))
  ) %>% 
  unnest(ttest_df)
```

Repeat the above for μ={1,2,3,4,5,6}

```{r}
sim_df2 = 
  expand_grid(
    mean = c(1, 2, 3, 4, 5, 6),
    iter = 1:5000
  ) %>% 
  mutate(
    ttest_df = map(.x = mean, ~ sim_ttest(mean = .x))
  ) %>% 
  unnest(ttest_df)
```

Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of μ on the x axis. Describe the association between effect size and power.

```{r}
sim_df2 %>% 
  mutate(
    reject = if_else(p_value < 0.05, "y", "n")) %>%
  group_by(mean, reject) %>%
  summarize(
    num_reject = n()) %>%
  mutate(
    prop_reject = num_reject/5000) %>%
  filter(reject == "y") %>%
  ggplot(
    aes(x = mean, y = prop_reject)) +
  geom_point() + geom_line()
```


Make a plot showing the average estimate of μ̂  on the y axis and the true value of μ on the x axis. Make a second plot (or overlay on the first) the average estimate of μ̂  only in samples for which the null was rejected on the y axis and the true value of μ on the x axis. Is the sample average of μ̂  across tests for which the null is rejected approximately equal to the true value of μ? Why or why not?

```{r}
mean_plot1 =
  sim_df2 %>% 
  ggplot(
    aes(x = mean, y = mu_hat)) +
  geom_point(alpha = 0.5) +
  labs(title = "All samples")

mean_plot2 =
  sim_df2 %>% 
    mutate(
    reject = if_else(p_value < 0.05, "y", "n")) %>%
  filter(reject == "y") %>%
  ggplot(
    aes(x = mean, y = mu_hat)) +
  geom_point(alpha = 0.5) +
  labs(title = "Samples for null rejected")

mean_plot1 + mean_plot2
```
